---
title: "Exponential Distribution"
author: "Dustin Lanning"
date: "April 4, 2019"
output:
  html_document: default
  pdf_document: default
---


##Overview

The purpose of the project is to investigate the exponential distribution and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of the exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Lambda will be set at 0.2 for all simulations. The investigation will compare the distribution of averages of 40 exponentials over 1,000 simulations.

##The Simulations

We set the variables for the simulations as follows:

```{r}
set.seed(108)
n<-40
lambda<-0.2
sims<-1000
```

Perform the simulations using the variables as follows:

```{r}
means = NULL
for (i in 1:sims) {
    means = c(means, mean(rexp(n, lambda)))
}
```

##Sample Mean vs. Theoretical Mean

###Sample Mean
Calculate the sample mean from the simulations.

```{r}
samMean<-mean(means)
```

###Theoretical Mean
Calculate the theoretical mean.

```{r}
theoMean<-1/lambda

meanComp<-samMean-theoMean
```

###Comparison
The sample mean of `r samMean` versus the theoretical mean of `r theoMean` for a difference of `r meanComp`.

##Sample Variance vs. Theoretical Variance

###Sample Variance
Calculate the sample variance from the simulations.

```{r}
samVar<-var(means)
```

###Theoretical Variance
Calculate the theoretical variance.

```{r}
theoVar<-(1/lambda)^2/n

varComp<-samVar-theoVar
```

###Comparison
The sample variance of `r samVar` versus the theoretical variance of `r theoVar` for a difference of `r varComp`.

##Distribution

The following is a density histogram of the simulated means. The density histogram has an overlay of a normal distribution with mean of 1/lambda and standard deviation of (1/lambda)*sqrt(n).

```{r}
library(ggplot2)
df<-data.frame(means)
ggplot(df, aes(x = means)) + geom_histogram(aes(y = ..density..), binwidth = 0.2, fill = "red", color = "black") + stat_function(geom = "line", fun = dnorm, args = list(mean = theoMean, sd = sd(means)), size = 1) + labs(x = "Simulated Means", y = "Density", title = "Density of Simulated Means")
```
